---
title: "Take Home Exercise 2"
---

# Introduction

#### In this study, we aim to undercover the spatio-temporal trends of COVID-19 vaccination in DKI Jakarta and the sub-districts to see their distribution across. A further analysis on how the distribution changed overtime will also be conducted.

# Data

#### The following data will be used in the analysis.

## Aspatial

##### Data Vaksinasi Berbasis Kelurahan (01 Juli 2021) \| Format: .xlsx \| Riwayat File Vaksinasi DKI Jakarta

## Geospatial

##### BATAS_DESA_DESEMBER_2019_DUKCAPIL_DKI_JAKARTA \| Format: Shapefile \| Indonesia Geospatial portal

# Install and load packages

```{r}
pacman::p_load(plyr , sf, tidyverse, readxl, tmap, maptools, kableExtra, sfdep, spdep, plotly, zoo)
```

# Data Import

### Geospatial Data

```{r}
jakarta_data <- st_read(dsn="data/geospatial",
                              layer="BATAS_DESA_DESEMBER_2019_DUKCAPIL_DKI_JAKARTA")
```

## Data Wrangling

##### Check for invalid geometries

##### *Reference from senior sample submissions for code for this section, with credit to Megan's [**Take-Home Exercise 1: Analysing and Visualising Spatio-temporal Patterns of COVID-19 in DKI Jakarta, Indonesia**](https://is415-msty.netlify.app/posts/2021-09-10-take-home-exercise-1/)*

```{r}
length(which(st_is_valid(jakarta_data) == FALSE))
```

#### No invalid geometries

##### Check for missing values

```{r}
jakarta_data[rowSums(is.na(jakarta_data))!=0,]
```

##### Remove missing fields

```{r}
jakarta_data <- na.omit(jakarta_data,c("DESA_KELUR"))
```

### Check coordinate system of data

```{r}
st_crs(jakarta_data)
```

#### Note: WGS84 is not appropriate, have to change to national coordinate system of Indonesia -- DGN95

```{r}
jakarta_data <- st_transform(jakarta_data, 23845)
```

##### Check if CRS changed

```{r}
st_crs(jakarta_data)
```

#### Removing redundant data

##### We are only interested in DKI Jakarta, hence we will be removing other islands or districts from the data.

##### Since KAB-KOTA which is 'City' will be the best way to filter them, we take a look at the cities in this dataset.

```{r}
unique(jakarta_data$"KAB_KOTA")
```

### Translation of column names for ease of handling data

##### Using rename() of dplyr package

```{r}
jakarta_data <- jakarta_data %>% 
  dplyr::rename(
    Object_ID=OBJECT_ID,
    Province=PROVINSI, 
    City=KAB_KOTA, 
    District=KECAMATAN, 
    Village_Code=KODE_DESA, 
    Village=DESA, 
    Sub_District=DESA_KELUR,
    Code=KODE, 
    Total_Population=JUMLAH_PEN
    )
```

#### Visualising to see all cities in this dataset

```{r}
tm_shape(jakarta_data) + 
  tm_polygons("Sub_District")
```

#### Removing redundant data

```{r}
jakarta_data <- filter(jakarta_data, City != "KEPULAUAN SERIBU")
```

##### Note: reduced from 267 observations to 261 observations

#### Retaining first 9 fields of jakarta_data as per assignment instructions

```{r}
jakarta_data <- jakarta_data %>% select(1:9)
```

##### Note: reduced to 10 variables

### Final visualisation of DKI Jakarta data

```{r}
tm_shape(jakarta_data) + 
  tm_polygons("Sub_District")
```

## Aspatial Data

## Aspatial Data Pre-processing Function

```{r}
# takes in an aspatial data filepath and returns a processed output
aspatial_preprocess <- function(filepath){
  # read xlsx file
  result_file <- read_xlsx(filepath)
  
  # Create the Date Column
  # the format of our files is: Vaccination DD MM YYYY Jarkarta
  # Starting Point: Vaccination
  # End Point: Jarkarta
  # Use [1] to indicate first element in the list
  # reference https://stackoverflow.com/questions/14249562/find-the-location-of-a-character-in-string
  startpoint <- gregexpr(pattern="Vaccination", filepath)[[1]] + 12
  endpoint <- gregexpr(pattern="Jarkarta", filepath)[[1]] - 2
  result_file$Date <- substr(filepath, startpoint, endpoint)
  
  # Retain the Relevant Columns
  result_file <- result_file %>% 
    select("Date", 
           "WILAYAH KOTA", 
           "KECAMATAN", 
           "KELURAHAN", 
           "SASARAN", 
           "BELUM VAKSIN")
  return(result_file)
}
```

## Feed files into preprocessing function

```{r}
# in the folder 'data/aspatial', find files with the extension '.xlsx' and add it to our fileslist 
# the full.names=TRUE prepends the directory path to the file names, giving a relative file path - otherwise, only the file names (not the paths) would be returned 
# reference: https://stat.ethz.ch/R-manual/R-devel/library/base/html/list.files.html
fileslist <-list.files(path = "data/aspatial", pattern = "*.xlsx", full.names=TRUE)

# afterwards, for every element in fileslist, apply aspatial_process function
dflist <- lapply(seq_along(fileslist), function(x) aspatial_preprocess(fileslist[x]))
```

## Convert dflist into actual dataframe

```{r}
cases_jakarta <- ldply(dflist, data.frame)
```

##### Check cases_jakarta

```{r}
glimpse(cases_jakarta)
```

## Format Date Column

```{r}
# parses the 'Date' column into Month(Full Name)-YYYY datetime objects
# reference: https://stackoverflow.com/questions/53380650/b-y-date-conversion-gives-na

# locale="ind" means that the locale has been set as Indonesia
Sys.setlocale(locale="ind")
```

```{r}
cases_jakarta$Date <- c(cases_jakarta$Date) %>% 
  as.Date(cases_jakarta$Date, format ="%d %B %Y")
glimpse(cases_jakarta)
```

## Renaming columns

```{r}
cases_jakarta <- cases_jakarta %>%
  dplyr::rename(
    Date=Date,
    City = WILAYAH.KOTA,
    District = KECAMATAN,
    Sub_District = KELURAHAN,
    Target_vaccine = SASARAN,
    Unvaccinated = BELUM.VAKSIN
  )
```

## Removing redundant cities

```{r}
cases_jakarta <- filter(cases_jakarta, City != "KAB.ADM.KEP.SERIBU")
```

## Data Integration of Geospatial and Aspatial

```{r}
# checks for column names of the dataframes
colnames(jakarta_data)
```

```{r}
colnames(cases_jakarta)
```

```{r}
# joins cases_jakarta to bd_jakarta based on Province, Sub_District and City
combined_jakarta <- left_join(jakarta_data, cases_jakarta,
                              by=c(
                                "District"="District", 
                                "Sub_District"="Sub_District",
                                "City"="City")
                              )
```

##### Now let's visualise subdistrict in terms of unvaccinated and target vaccine

```{r}
target_vaccine <- tm_shape(combined_jakarta) +
  tm_borders(alpha = 0.5) +
  tm_fill("Target_vaccine")

not_vaccinated <- tm_shape(combined_jakarta) +
  tm_borders(alpha = 0.5) +
  tm_fill("Unvaccinated")

tmap_arrange(target_vaccine, not_vaccinated)
```

#### As shown above, there are grey areas in the map. These are missing values which might be caused by mismatched records from joining subdistricts/city from the 2 dataframes.

## Correcting mismatched records

#### To find, we see what was in jakarta_data but not in cases_jakarta

```{r}
vaccine_subdistrict <- c(cases_jakarta$Sub_District)
jakarta_subdistrict <- c(jakarta_data$Sub_District)

unique(vaccine_subdistrict[!(vaccine_subdistrict %in% jakarta_subdistrict)])
```

```{r}
unique(jakarta_subdistrict[!(jakarta_subdistrict %in% vaccine_subdistrict)])
```

### Let us visualise the mismatched records

| Geospatial Data (jakarta_subdistrict) | Aspatial Data (vaccine_subdistrict) |
|------------------------------------|------------------------------------|
| BALEKAMBANG                           | BALE KAMBANG                        |
| HALIM PERDANA KUSUMA                  | HALIM PERDANA KUSUMAH               |
| JATIPULO                              | JATI PULO                           |
| TENGAH                                | KAMPUNG TENGAH                      |
| KRENDANG                              | KERENDANG                           |
| KRAMATJATI                            | KRAMAT JATI                         |
| PALMERIAN                             | PAL MERIAN                          |
| PINANGRANTI                           | PINANG RANTI                        |
| RAWAJATI                              | RAWA JATI                           |

## Renaming mismatched records

```{r}
# where bd_jakarta is a mismatched value, replace with the correct value
jakarta_data$Sub_District[jakarta_data$Sub_District == 'BALEKAMBANG'] <- 'BALE KAMBANG'
jakarta_data$Sub_District[jakarta_data$Sub_District == 'HALIM PERDANA KUSUMA'] <- 'HALIM PERDANA KUSUMAH'
jakarta_data$Sub_District[jakarta_data$Sub_District == 'JATIPULO'] <- 'JATI PULO'
jakarta_data$Sub_District[jakarta_data$Sub_District == 'TENGAH'] <- 'KAMPUNG TENGAH'
jakarta_data$Sub_District[jakarta_data$Sub_District == 'KRAMATJATI'] <- 'KRAMAT JATI'
jakarta_data$Sub_District[jakarta_data$Sub_District == 'KRENDANG'] <- 'KERENDANG'
jakarta_data$Sub_District[jakarta_data$Sub_District == 'PALMERIAM'] <- 'PAL MERIAM'
jakarta_data$Sub_District[jakarta_data$Sub_District == 'PINANGRANTI'] <- 'PINANG RANTI'
jakarta_data$Sub_District[jakarta_data$Sub_District == 'RAWAJATI'] <- 'RAWA JATI'
```

## Joining both dataframes by sub-district

```{r}
combined_jakarta <- left_join(jakarta_data, cases_jakarta,
                          by=c("Sub_District" = "Sub_District"))
```

## Visualising data once again

```{r}
target_vaccine <- tm_shape(combined_jakarta) +
  tm_borders(alpha = 0.5) +
  tm_fill("Target_vaccine")

not_vaccinated <- tm_shape(combined_jakarta) +
  tm_borders(alpha = 0.5) +
  tm_fill("Unvaccinated")

tmap_arrange(target_vaccine, not_vaccinated)
```

## Calculations

### Computation of Monthly Vaccination Rate

##### Here, we subtract the quantity of non-vaccinated to subtract from the number of Target Vaccine, and divide it to retrieve a percentage decimal of vaccination rate.

```{r}
vaccine_rate <- cases_jakarta %>%
  inner_join(jakarta_data, by=c("Sub_District" = "Sub_District")) %>%
  group_by(Sub_District, Date) %>%
  summarise(`vaccination_rate` = (Target_vaccine - Unvaccinated)/Target_vaccine)
```

```{r}
glimpse(vaccine_rate)
```

```{r}
vaccine_rate_pivot <- vaccine_rate %>% ungroup() %>% pivot_wider(names_from = Date, values_from = vaccination_rate)

```

## Mapping Monthly Vaccination Rate

#### Convert to sf dataframe first

```{r}
combined_jakarta <- st_as_sf(combined_jakarta)
```

```{r}
vaccine_rate_pivot <- vaccine_rate_pivot %>% 
  left_join(jakarta_data, by=c("Sub_District" = "Sub_District"))
vaccine_rate_pivot <- st_as_sf(vaccine_rate_pivot)
```

### Create map helper function for recursive map plotting

```{r}
map_function <- function(df, varname) {
tm_shape(df) +
  tm_fill(varname,
          style = "quantile",
          palette = "Blues",
          title = "Vaccination Rate") +
  tm_layout(main.title = varname,
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type = "8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha = 0.2)
}
```

### Executing the function

```{r}
tmap_mode("plot")
tmap_arrange(map_function(vaccine_rate_pivot, "2021-07-01"),
             map_function(vaccine_rate_pivot, "2021-08-01"),
             map_function(vaccine_rate_pivot, "2021-09-01"),
             map_function(vaccine_rate_pivot, "2021-10-01"))
```

```{r}
tmap_arrange(map_function(vaccine_rate_pivot, "2021-11-01"),
             map_function(vaccine_rate_pivot, "2021-12-01"),
             map_function(vaccine_rate_pivot, "2022-01-01"),
             map_function(vaccine_rate_pivot, "2022-02-01"))
```

```{r}
tmap_arrange(map_function(vaccine_rate_pivot, "2022-03-01"),
             map_function(vaccine_rate_pivot, "2022-04-01"),
             map_function(vaccine_rate_pivot, "2022-05-01"),
             map_function(vaccine_rate_pivot, "2022-06-01"))
```

## Spatial Pattern Analysis

#### As the palette darkens to a darker shade of blue, it shows a higher vaccination rate. With a varying shade of blue across the entire Jakarta region, it is observed that the vaccines administered were inconsistent -- some sub-districts with higher vaccination rates in the first few months of the study period did not maintain as high during later parts of the study period. This could be due to the fact that their citizens in that area have been vaccinated and hence, vaccination efforts are stronger in other sub-districts subsequently.

# 10. Local GI\* Analysis

#### We will be conducting hotspot and coldspot analysis using G-statistics.

#### The steps to this analysis are:

#### 1) Derive spatial weight matrix

#### 2) Compute Gi statistics

#### 3) Mapping Gi statistics

## Derive contiguity weights using Queen's method (sfdep() package)

```{r}
set.seed(1234)
```

```{r}
wm_idw <- vaccine_rate_pivot %>%
  mutate(nb = st_contiguity(geometry),
         wts = st_inverse_distance(nb, geometry, scale = 1, alpha = 1),
         .before = 1)
```

## Compute Gi\* values of vaccination rates for each month

```{r}
HCSA_july2021 <- wm_idw %>% 
  mutate(local_Gi = local_gstar_perm(
    `2021-07-01`, nb, wt, nsim = 99),
         .before = 1) %>%
  unnest(local_Gi)

HCSA_aug2021 <- wm_idw %>% 
  mutate(local_Gi = local_gstar_perm(
    `2021-08-01`, nb, wt, nsim = 99),
         .before = 1) %>%
  unnest(local_Gi)

HCSA_sept2021 <- wm_idw %>% 
  mutate(local_Gi = local_gstar_perm(
    `2021-09-01`, nb, wt, nsim = 99),
         .before = 1) %>%
  unnest(local_Gi)

HCSA_oct2021 <- wm_idw %>% 
  mutate(local_Gi = local_gstar_perm(
    `2021-10-01`, nb, wt, nsim = 99),
         .before = 1) %>%
  unnest(local_Gi)

HCSA_nov2021 <- wm_idw %>% 
  mutate(local_Gi = local_gstar_perm(
    `2021-11-01`, nb, wt, nsim = 99),
         .before = 1) %>%
  unnest(local_Gi)

HCSA_dec2021 <- wm_idw %>% 
  mutate(local_Gi = local_gstar_perm(
    `2021-12-01`, nb, wt, nsim = 99),
         .before = 1) %>%
  unnest(local_Gi)

HCSA_jan2022 <- wm_idw %>% 
  mutate(local_Gi = local_gstar_perm(
    `2022-01-01`, nb, wt, nsim = 99),
         .before = 1) %>%
  unnest(local_Gi)

HCSA_feb2022 <- wm_idw %>% 
  mutate(local_Gi = local_gstar_perm(
    `2022-02-01`, nb, wt, nsim = 99),
         .before = 1) %>%
  unnest(local_Gi)

HCSA_mar2022 <- wm_idw %>% 
  mutate(local_Gi = local_gstar_perm(
    `2022-03-01`, nb, wt, nsim = 99),
         .before = 1) %>%
  unnest(local_Gi)

HCSA_april2022 <- wm_idw %>% 
  mutate(local_Gi = local_gstar_perm(
    `2022-04-01`, nb, wt, nsim = 99),
         .before = 1) %>%
  unnest(local_Gi)

HCSA_may2022 <- wm_idw %>% 
  mutate(local_Gi = local_gstar_perm(
    `2022-05-01`, nb, wt, nsim = 99),
         .before = 1) %>%
  unnest(local_Gi)

HCSA_june2022 <- wm_idw %>% 
  mutate(local_Gi = local_gstar_perm(
    `2022-06-01`, nb, wt, nsim = 99),
         .before = 1) %>%
  unnest(local_Gi)
```

## Create map objects of Gi\* map, then plot the map.

#### For each map, only hot and cold spot areas are displyed. (p-value less than 0.05)

```{r}
#July 2021
HCSA_july2021_sig <- HCSA_july2021  %>%
  filter(p_sim < 0.05)

HCSA_july2021_map <- tm_shape(HCSA_july2021) +
                      tm_polygons() +
                      tm_borders(alpha = 0.5) +
                    tm_shape(HCSA_july2021_sig) +
                      tm_fill("gi_star") + 
                      tm_borders(alpha = 0.4) + 
                      tm_layout(main.title = "July 2021")

#Aug 2021
HCSA_aug2021_sig <- HCSA_aug2021  %>%
  filter(p_sim < 0.05)

HCSA_aug2021_map <-tm_shape(HCSA_aug2021) +
                      tm_polygons() +
                      tm_borders(alpha = 0.5) +
                    tm_shape(HCSA_aug2021_sig) +
                      tm_fill("gi_star") + 
                      tm_borders(alpha = 0.4) + 
                      tm_layout(main.title = "Aug 2021")

#Sept 2021
HCSA_sept2021_sig <- HCSA_sept2021  %>%
  filter(p_sim < 0.05)

HCSA_sept2021_map <- tm_shape(HCSA_sept2021) +
                      tm_polygons() +
                      tm_borders(alpha = 0.5) +
                    tm_shape(HCSA_sept2021_sig) +
                      tm_fill("gi_star") + 
                      tm_borders(alpha = 0.4) + 
                      tm_layout(main.title = "Sept 2021")

#Oct 2021
HCSA_oct2021_sig <- HCSA_oct2021  %>%
  filter(p_sim < 0.05)

HCSA_oct2021_map <-tm_shape(HCSA_oct2021) +
                    tm_polygons() +
                    tm_borders(alpha = 0.5) +
                  tm_shape(HCSA_oct2021_sig) +
                    tm_fill("gi_star") + 
                    tm_borders(alpha = 0.4) + 
                      tm_layout(main.title = "Oct 2021")

# Nov 2021
HCSA_nov2021_sig <- HCSA_nov2021  %>%
  filter(p_sim < 0.05)

HCSA_nov2021_map <-tm_shape(HCSA_nov2021) +
                    tm_polygons() +
                    tm_borders(alpha = 0.5) +
                  tm_shape(HCSA_nov2021_sig) +
                    tm_fill("gi_star") + 
                    tm_borders(alpha = 0.4) + 
                      tm_layout(main.title = "Nov 2021")
# Dec 2021
HCSA_dec2021_sig <- HCSA_dec2021  %>%
  filter(p_sim < 0.05)

HCSA_dec2021_map <-tm_shape(HCSA_dec2021) +
                  tm_polygons() +
                  tm_borders(alpha = 0.5) +
                tm_shape(HCSA_dec2021_sig) +
                  tm_fill("gi_star") + 
                  tm_borders(alpha = 0.4) + 
                      tm_layout(main.title = "Dec 2021")

# Jan 2022
HCSA_jan2022_sig <- HCSA_jan2022  %>%
  filter(p_sim < 0.05)

HCSA_jan2022_map <- tm_shape(HCSA_jan2022) +
                    tm_polygons() +
                    tm_borders(alpha = 0.5) +
                  tm_shape(HCSA_jan2022_sig) +
                    tm_fill("gi_star") + 
                    tm_borders(alpha = 0.4) + 
                      tm_layout(main.title = "Jan 2022")

# Fen 2022
HCSA_feb2022_sig <- HCSA_feb2022  %>%
  filter(p_sim < 0.05)

HCSA_feb_2022_map <- tm_shape(HCSA_feb2022) +
                      tm_polygons() +
                      tm_borders(alpha = 0.5) +
                    tm_shape(HCSA_feb2022_sig) +
                      tm_fill("gi_star") + 
                      tm_borders(alpha = 0.4) + 
                      tm_layout(main.title = "Feb 2022")

# Mar 2022
HCSA_mar2022_sig <- HCSA_mar2022  %>%
  filter(p_sim < 0.05)

HCSA_mar2022_map <- tm_shape(HCSA_mar2022) +
                      tm_polygons() +
                      tm_borders(alpha = 0.5) +
                    tm_shape(HCSA_mar2022_sig) +
                      tm_fill("gi_star") + 
                      tm_borders(alpha = 0.4) + 
                      tm_layout(main.title = "Mar 2022")

#April 2022
HCSA_april2022_sig <- HCSA_april2022  %>%
  filter(p_sim < 0.05)

HCSA_april2022_map <- tm_shape(HCSA_april2022) +
                        tm_polygons() +
                        tm_borders(alpha = 0.5) +
                      tm_shape(HCSA_april2022_sig) +
                        tm_fill("gi_star") + 
                        tm_borders(alpha = 0.4) + 
                      tm_layout(main.title = "April 2022")

# May 2022
HCSA_may2022_sig <- HCSA_may2022  %>%
  filter(p_sim < 0.05)

HCSA_may2022_map <- tm_shape(HCSA_may2022) +
                      tm_polygons() +
                      tm_borders(alpha = 0.5) +
                    tm_shape(HCSA_may2022_sig) +
                      tm_fill("gi_star") + 
                      tm_borders(alpha = 0.4) + 
                      tm_layout(main.title = "May 2022")

# June 2022
HCSA_june2022_sig <- HCSA_june2022  %>%
  filter(p_sim < 0.05)

HCSA_june2022_map <- tm_shape(HCSA_june2022) +
                      tm_polygons() +
                      tm_borders(alpha = 0.5) +
                    tm_shape(HCSA_june2022_sig) +
                      tm_fill("gi_star") + 
                      tm_borders(alpha = 0.4) + 
                      tm_layout(main.title = "Jun 2022")
```

```{r}
tmap_arrange(HCSA_july2021_map, HCSA_aug2021_map, HCSA_sept2021_map, HCSA_oct2021_map)
```

```{r}
tmap_arrange(HCSA_mar2022_map, HCSA_april2022_map, HCSA_may2022_map, HCSA_june2022_map)
```

# Conclusion

#### Hotspot areas shaded in green indicates subdistricts with high vaccination rates and coldspot areas indcates subdistricts with low vaccination rates. Hence, we can conclude there is an initial major hotspot.

## Emerging Hotspot Analysis (EHSA)

#### We will be conducting an EHSA on the monthly vaccination rates to see how hot and coldspots evolve over time.

### Build a time series cube

##### spacetime() of sfdep will be used to create a spatio-temporal cube

```{r}
vaccine_rate_st <- spacetime(vaccine_rate, jakarta_data,
                             .loc_col = "Sub_District",
                             .time_col = "Date")

vaccine_rate_st
```

```{r}
is_spacetime_cube(vaccine_rate_st)
```

#### The TRUE return confirms that vaccine_rate_st object is indeed an time-space cube.

## Deriving Spatial Weights

```{r}
# vaccine_nb <- vaccine_rate_st %>%
#   activate("geometry") %>%
#   mutate(nb = include_self(st_contiguity(geometry)),
#          wt = st_inverse_distance(nb, geometry,
#                                   scale = 1,
#                                   alpha = 1),
#          .before = 1)%>%
#   set_wts("wt") %>%
#   set_nbs("nb")
# 
# head(vaccine_nb)
```

## Computing Local Gi\*

```{r}
# gi_stars <- vaccine_nb %>% 
#   group_by(`Date`) %>% 
#   mutate(gi_star = local_gstar_perm(
#     vaccine_rate, nb, wt)) %>% 
#   tidyr::unnest(gi_star)
# 
# gi_stars
```

## Man Kendall Test

```{r}
june_rate <- vaccine_rate %>%
  filter(Date == as.Date("2022-06-01"))%>%
  arrange(desc(`vaccination_rate`))

head(june_rate)
```

#### The 3 sub-districts with the highest vaccination rates are: HALIM PERDANA KUSUMAH, SRENGSENG SAWAH and MANGGARAI SELATAN.

## Observing these 3 sub-districts for Man Kendall Test

```{r}
#HALIM PERDANA KUSUMAH

# cbg_hpk <- gi_stars %>% 
#   ungroup() %>% 
#   filter(Sub_District == "HALIM PERDANA KUSUMAH") |> 
#   select(Sub_District, Date, gi_star)
# 
# ggplot(data = cbg_hpk, 
#        aes(x = Date, 
#            y = gi_star)) +
#   geom_line() +
#   theme_light()
```

```{r}
#SRENGSENG SAWAH

# cbg_ss <- gi_stars %>% 
#   ungroup() %>% 
#   filter(Sub_District == "SRENGSENG SAWAH") |> 
#   select(Sub_District, Date, gi_star)
# 
# ggplot(data = cbg_ss, 
#        aes(x = Date, 
#            y = gi_star)) +
#   geom_line() +
#   theme_light()
```

```{r}
#MANGGARAI SELATAN

# cbg_ms <- gi_stars %>% 
#   ungroup() %>% 
#   filter(Sub_District == "MANGGARAI SELATAN") |> 
#   select(Sub_District, Date, gi_star)
# 
# ggplot(data = cbg_ms, 
#        aes(x = Date, 
#            y = gi_star)) +
#   geom_line() +
#   theme_light()
```
